---
title: "Boruta feature selection for treebased methods"
author: "Johannes Fjeldså"
date: "2024-03-05"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(Boruta)
library(dplyr)
library(randomForest)
library(purrr)
library(doParallel)
library(caret)
library(xgboost)
```

```{r}
scenario_indx_key <- list('ssp126' = 0, 'ssp585' = 1)
years <- seq(2015, 2060, 5)
#years <- seq(2035, 2040, 1)

cross_section_table_dir <- "D:/Programmering/msc/Masterthesis_S23 DataFiles/unscaled"
scaled_cross_sections <- list()
for (year in years) {
   file_path <- file.path(cross_section_table_dir, paste0("cross_section_", year, ".csv"))#, "_standscaled.csv"))
   df <- read.csv(file_path)
   df <- df %>%
         filter(scenario %in% names(scenario_indx_key)) 
   matches <- match(df$scenario, names(scenario_indx_key))
   df$scenario_indx <- unlist(scenario_indx_key)[matches]
   scaled_cross_sections[[year]] <- df
}
```


# Performe Boruta feature selection
```{r}
run_boruta <- function(data, model_name, model_params=NULL) {
  
  # Set estimator
  if (model_name == 'RF'){
    getImp <- getImpRfZ
  } else if (model_name == 'XGB'){
    getImp <- getImpXgboost#Normalized
  } else if (model_name == 'RFe'){
    getImp <- getImpFerns
  } else {
    stop('Model not recognized')
  }
  # Set default parameters
  boruta_params <- list(formula = scenario_indx ~ .-scenario, 
                        data = data, 
                        doTrace = 0, 
                        maxRuns = 100,
                        getImp=getImp)
  
  if (!is.null(model_params)){
    boruta_params <- c(boruta_params, model_params)
  }
    

  boruta_output <- do.call(Boruta, boruta_params)
                   
  final_boruta <- TentativeRoughFix(boruta_output)
  boruta_df <- as.data.frame(attStats(final_boruta))
  boruta_df <- boruta_df[order(as.vector(rownames(boruta_df))), ]
  return(list(boruta_df = boruta_df, boruta_output = boruta_output))
}
```

```{r}
extract_results <- function(boruta_dataframes, years) {
  
  new_colnames <- rownames(boruta_dataframes[[2015]])
  new_colnames <- c('year', new_colnames)
  boruta_scores_df <- data.frame(matrix(ncol = length(new_colnames), nrow = length(years)))
  colnames(boruta_scores_df) <- new_colnames
  boruta_decisions_df <- data.frame(matrix(ncol = length(new_colnames), nrow = length(years)))
  colnames(boruta_decisions_df) <- new_colnames
  
  i=1
  for (year in years) {
    importance <- boruta_dataframes[[year]]$medianImp
    importance <- c(year, importance)
    boruta_scores_df[i, ] <- importance
    decision <- boruta_dataframes[[year]]$decision
    decision <- c(year, decision)
    boruta_decisions_df[i, ] <- decision
    i = i + 1
  }

write.csv(boruta_decisions_df, "C:/Users/fjejo/Downloads/boruta_decisions_df.csv")
write.csv(boruta_scores_df, "C:/Users/fjejo/Downloads/boruta_scores_df.csv")

}
```

Now we can run the full boruta alg. for every year and extract the results. 

## Run Boruta with RF

```{r, fig.width=10, fig.height=8}

boruta_dataframes = list()
indx = 0
for (year in years) {
  
  cross_section_df <- scaled_cross_sections[[year]]

  boruta_results <- run_boruta(cross_section_df, 'RF')
  boruta_dataframes[[year]] <- boruta_results$boruta_df
  boruta_output <- boruta_results$boruta_output
  
  if (year%%5==0) {
    par(mar = c(12, 4, 4, 1) + 1.5)  # Adjust the margin to make room for tick labels
    plot(boruta_output, cex.axis=.7, las=2, xlab="", main=paste("Variable Importance in cross section (RF)", year))    }
}

extract_results(boruta_dataframes, years)
```
## Tune XGB

# XGB

## Tune xgboost

We need one for each year since the data is different for each year. We start by creating a function that makes an initial tuning with grio search. I will use the following parameters:

-   nrounds: Represents the number of boosting rounds during training.

-   max_depth: Controls the maximum depth of each tree in the ensemble, affecting model complexity and overfitting.

-   eta: Also known as learning rate, determines the step size at each iteration.

-   gamma: Governs the minimum loss reduction required to make a further partition on a leaf node.

-   colsample_bytree: Specifies the fraction of columns to be subsampled when constructing each tree.

-   min_child_weight: Defines the minimum sum of instance weight needed in a child node.

-   subsample: Represents the fraction of observations to be sampled for each tree.

The function uses cross validation to find the best parameters and returns the best model for each cross section year. Based on code from [amirali_n's Github](https://amirali-n.github.io/ExtremeGradientBoosting/).

```{r}
tune_xgb <- function(data, X_colnames, y_colname=NaN, 
                     nrounds = NaN, 
                     max_depth = NaN, 
                     eta = NaN, 
                     gamma = NaN, 
                     colsample_bytree = NaN, 
                     min_child_weight = NaN, 
                     subsample = NaN) {
  
  # make it run faster (in parallell)
  myCl=makeCluster(detectCores()-1)
  registerDoParallel(myCl)
  
  # set up the grid parameters
  nrounds <- ifelse(is.nan(nrounds), 100, nrounds)
  max_depth <- ifelse(is.nan(max_depth), 5, max_depth)
  eta <- ifelse(is.nan(eta), 0.3, eta)
  gamma <- ifelse(is.nan(gamma), 0, gamma)
  colsample_bytree <- ifelse(is.nan(colsample_bytree), 1, colsample_bytree)
  min_child_weight <- ifelse(is.nan(min_child_weight), 1, min_child_weight)
  subsample <- ifelse(is.nan(subsample), 1, subsample)
 
  
  
  xgb_grid <- expand.grid(nrounds=nrounds,
                          eta=eta,
                          max_depth=max_depth,
                          gamma=gamma,
                          subsample = subsample,
                          min_child_weight = min_child_weight, 
                          colsample_bytree = colsample_bytree)
  # Set rules for actual training
  xgb_control <- trainControl(method = "cv",
                              number = 3,
                              verboseIter = TRUE,
                              returnData = FALSE,
                              returnResamp = "none",
                              classProbs = TRUE,
                              allowParallel = TRUE)
  
  
  xgb_train <- train(x = data[, X_colnames],
                     y = data[['scenario']], 
                     trControl = xgb_control,
                     tuneGrid = xgb_grid,
                     method = "xgbTree")
  
  # close parallellcluster
  stopCluster(myCl)
  
  best_params <- list("eta"=xgb_train$bestTune$eta,
                 "max_depth"=xgb_train$bestTune$max_depth,
                 "gamma"=xgb_train$bestTune$gamma,
                 "min_child_weight"=xgb_train$bestTune$min_child_weight,
                 "nthread"=4,
                 "objective"="binary:logistic")
  
  xgb_crv = xgb.cv(params = best_params,
                   data = as.matrix(data[, X_colnames]),
                   nrounds = 500,
                   nfold = 3,
                   label = data[['scenario_indx']],
                   showsd = TRUE,
                   metrics = "auc",
                   stratified = TRUE,
                   verbose = FALSE,
                   print_every_n = 1L,
                   early_stopping_rounds = 50)
  
  best_params["nrounds"] <- xgb_crv$best_iteration
  
  return(best_params)
}
```


```{r, warning=FALSE, message=FALSE}
features <- names(scaled_cross_sections[[2035]])[!names(scaled_cross_sections[[2035]]) %in% c('scenario', 'scenario_indx')]
last_year = years[length(years)]

xgb_models <- list()
for (year in years) {
  print(paste(year, 'of', last_year))
  cross_section_df <- scaled_cross_sections[[year]]
  best_params <- tune_xgb(cross_section_df, X_colnames=features, 
                          nrounds=100, 
                          max_depth=c(3, 5, 10), 
                          eta=c(0.0001, 0.001, 0.01, 0.1, 0.3, 0.5, 0.7, 1), 
                          gamma=0, 
                          colsample_bytree=seq(0.5, 1, 0.1), 
                          min_child_weight=c(1, 3, 5), 
                          subsample=.7)
  
  xgb_models[[year]] <- best_params
}
```

We now have a rough idea of what our xgboost model will look like when it is performing the actual training. We now apply the boruta alg. to every year using the corresponding xgboost model as the estimator. This will return the important features for each year assuming one uses xgboost for the classification.

```{r}
getImpXgboostNormalized <- function(x, y, nrounds = 5, verbose = 0, ...) {
 # Call the original getImpXgboost function
 imp_scores <- Boruta::getImpXgboost(x, y, nrounds = nrounds, verbose = verbose, ...)
  
 # Normalize the importance scores
 # Here, you would implement your normalization method
 # For demonstration, let's use min-max normalization
 normalized_scores <- (imp_scores - min(imp_scores)) / (max(imp_scores) - min(imp_scores))
  
 # Return the normalized scores
 return(normalized_scores)
}
```

```{r, fig.width=10, fig.height=8}

boruta_dataframes = list()
indx = 0
for (year in years) {
  xgb_models[[year]] -> best_params
  cross_section_df <- scaled_cross_sections[[year]]

  boruta_results <- run_boruta(cross_section_df, 'XGB', best_params)
  boruta_dataframes[[year]] <- boruta_results$boruta_df
  boruta_output <- boruta_results$boruta_output
  
  if (year%%1==0) {
    par(mar = c(12, 4, 4, 1) + 1.5)  # Adjust the margin to make room for tick labels
    plot(boruta_output, cex.axis=.7, las=2, xlab="", main=paste("Variable Importance in cross section (XGB)", year))    }
}

extract_results(boruta_dataframes, years)


```
## Check performance of features

```{r}
library(tidymodels)
```

```{r}
df <- scaled_cross_sections[[2040]]

set.seed(3)
gap_split <- initial_split(df, prop = 0.75)
training_data <- training(gap_split)
test_data <- testing(gap_split)

best_params <- tune_xgb(training_data, X_colnames=features, 
                        nrounds=100, 
                        max_depth=c(3, 5, 10), 
                        eta=seq(0.1, 1, 0.2), 
                        gamma=0, 
                        colsample_bytree=1, 
                        min_child_weight=c(1, 3, 5), 
                        subsample=.7)
```
```{r, warning=FALSE, message=FALSE}
boruta_features =  c('fdETCCDI..nomask', 'pr..nomask', 'pr..sea_mask', 'tas..land_mask', 'tas..nomask', 'tas..sea_mask', 'txxETCCDI..nomask', 'txxETCCDI..sea_mask')
nomask_features = c('fdETCCDI..nomask', 'gslETCCDI..nomask', 'pr..nomask', 'tas..nomask', 'txxETCCDI..nomask')
all_features = names(scaled_cross_sections[[2030]])[!names(scaled_cross_sections[[2030]]) %in% c('scenario', 'scenario_indx')]
mrmr_boruta_features = c('tas..sea_mask', 'tas..nomask', 'txxETCCDI..nomask', 'txxETCCDI..sea_mask', 'tas..land_mask', 'pr..nomask')

maske_names = c('boruta', 'nomask', 'all', 'mRMR')
feature_selections = c(boruta_features, nomask_features, all_features, mrmr_boruta_features)

df <- scaled_cross_sections[[2025]]

accuracy <- list(boruta = numeric(10), 
                 nomask = numeric(10), 
                 all = numeric(10), 
                 mRMR = numeric(10))

for (seed in 1:100) {
  set.seed(seed)
  print(seed)
  train_indices <- createDataPartition(df[['scenario_indx']], p = 0.8, list = FALSE)
  train_data <- df[train_indices, ]
  test_data <- df[-train_indices, ]
  
  for (k in seq_along(maske_names)) {
    name = maske_names[k]
    feature_selection = feature_selections[k]
    
    dtrain <- xgb.DMatrix(data = as.matrix(train_data[, feature_selection]), label = train_data$scenario_indx)
    model <- xgb.train(params = best_params, data = dtrain, nrounds = 100)
    
    dtest <- xgb.DMatrix(data = as.matrix(test_data[, feature_selection]), label = test_data$scenario_indx)
    preds <- predict(model, dtest)
    # Convert predictions to class labels
    pred_class <- ifelse(preds > 0.5, "1", "0") # Assuming binary classification
    
    # Create a confusion matrix
    accuracy[[name]][seed] <- confusionMatrix(table(pred_class, test_data$scenario_indx))$overall[1]
  }
}
print(accuracy)

```

```{r}
print(accuracy)
boxplot(accuracy, names = c("Boruta", "Nomask", "All", "mRMR"), ylab = "Values")
```





